Manuscript 1 code

### 2024-10-25 manuscript_1_code.Rmd
cleaned code up to figure 5. Moved figure 2 to supplements. moved diversity-synhcrony plot to supplements.

## Figure 2a Richness @ Quad by Year

### plot
```{r figure 2a}
(figure_2a <- ggplot() +
  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
             position = position_dodge(0.4)) +
  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
                                             colour = habitat), width = 0, size = 1.3, position = position_dodge(0.4)) +
  scale_colour_manual(values = habitat_colours) +
  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
   labs(y = "Macroalgal Richness",
        x = "Year",
        title = "") +
   trends_theme)

# ggsave(filename = "output/figure_2a.png", figure_2a, height = 10, width = 14)
```

### plot
```{r plot figure 2b}
(figure_2b <- ggplot() +
  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
             position = position_dodge(0.4)) +
#  geom_jitter(data = alpha_diversity_site_macro, 
#  aes(x = year, y = richness, colour = habitat, alpha = 0.05), size = 2, width = 0.4, height = 0) +
  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0,  position = position_dodge(0.4)) +
  scale_colour_manual(values = habitat_colours) +
  annotate("segment", x = 2006, xend = 2010, y = 12, yend = 12, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_richness_df$year)) +
   labs(y = "",
        x = "Year",
        title = "") +
  trends_theme)

# ggsave(filename = "output/figure_2b.png", figure_2b, height = 10, width = 14)
```

## Figure 2 Full Plot
```{r ggpubr fig2, fig.height = 10, fig.width=22}
fig2_plot_list <- list(figure_2a, figure_2b)
(figure_2 <- ggarrange(figure_2a, figure_2b,
                      ncol = 2, nrow = 1, 
                      labels = c("a.", "b."),
                      common.legend = TRUE,
                      legend = "bottom", 
                      font.label = list(size = 24, color = "black", face = "plain")))

# ggsave(filename = "output/figure_2v4.png", figure_2, height = 10, width = 22)
```

# Figure 3 - ecosystem function

@@

diversity_stability_site <- alpha_diversity_site_macro %>%
  dplyr::group_by(site_habitat, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            shannon_mean = mean(shannon), 
            cover_mean = mean(prop_cover),
            cover_cv = CV(prop_cover),
            cover_stability = 1/CV(prop_cover))

@@

dss_ranges_site <- extract_ranges(diversity_stability_synchrony_site, "habitat",
                                  c("richness_mean", "cover_stability", "synchrony"))

@@

filtered_rich_stab_effects <- rich_stab_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)
```

@@

#### no habitat

```{r richness stability no hab}
# rich_stab_mod_2 <- glmmTMB(cover_stability~ richness_mean + (1|site), family = Gamma(link = "log"), data = # diversity_stability_synchrony)
# summary(rich_stab_mod_2)
# hist(residuals(rich_stab_mod_2)) # a chunkier version
# plot(residuals(rich_stab_mod_2) ~ fitted(rich_stab_mod_2)) # heteroskedastic - bring to group
# r.squaredGLMM(rich_stab_mod_2) # 0.750, 0.757

```{r 4a no hab setup}
# rich_stab_emm_2 <-as.data.frame(emmeans(rich_stab_mod_2, ~richness_mean,
#                                      at = list(richness_mean = seq(min(diversity_stability_synchrony$richness_mean),
#                                                             max(diversity_stability_synchrony$richness_mean),
#                                                             0.01)))) 
# 
```

#### no habitat plot

```{r figure 4a no habitat}
# (figure_4a_no_hab <- 
#   ggplot() +
#   geom_point(data = diversity_stability_synchrony,
#              aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
#   geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
#   geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
#               alpha = 0.3) +
#   geom_line(data = rich_stab_emm_2, aes(x = richness_mean, y = exp(emmean)), size = 2, colour = "black") +
#   geom_ribbon(data = rich_stab_emm_2, aes(x = richness_mean, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL)), fill = "black",
#               alpha = 0.3) +
#   scale_colour_manual(values = habitat_colours) +
#   scale_fill_manual(values = habitat_colours) +
#   labs(y = "Stability\n", x = "Macroalgal Richness") +
#   theme_bw() + 
#   theme(axis.title.x = element_text(size=18), 
#         axis.title.y = element_text(size=18), 
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_rect(fill = "white"),
#         legend.text = element_text(size = rel(1)), # increase legend text
#         legend.title = element_blank(),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal', # make legend horizontal
#         legend.box.background = element_rect(colour = "black"))  # add box around legend
#          )
# 
# ggsave(filename = "output/figure_4a_with_no_hab.png", figure_4a_no_hab, height = 10, width = 14)
```

## Figure 4b - functional group richness stability
### Setup
```{r functional richness through time at quad level}
fg_summary_wide_meta <- merge(fg_summary_wide, meta_df, by = c("location", "site", "habitat", "year") )
fg_summary_wide_meta_through_time <- fg_summary_wide_meta %>%
  group_by(location, habitat, site) %>%
  summarise(fg_richness_mean = mean(fg_richness_quad),
            fg_richness_se = sd(fg_richness_quad)/n())

#merge with fg data 
fg_stability_quad_level <- merge(fg_summary_wide_meta_through_time, diversity_stability_synchrony, 
                              by = 
                                c("location","site", "habitat"))

dss_ranges_fg <- fg_stability_quad_level %>%
  group_by(habitat) %>%
  summarise(min_richness_fg = min(fg_richness_mean),
            max_richness_fg = max(fg_richness_mean),
            min_stability= min(cover_stability, na.rm = TRUE),
            max_stability = max(cover_stability, na.rm = TRUE))
```

### figure 4b model
```{r fig4b fg richness stability mod}

rich_stab_mod_fg <- glmmTMB(cover_stability ~ fg_richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = fg_stability_quad_level)

@@

hist(residuals(rich_stab_mod_fg)) # looks fine
plot(residuals(rich_stab_mod_fg) ~ fitted(rich_stab_mod_fg)) # same as usual 
r.squaredGLMM(rich_stab_mod_fg) # 0.819, 0.828
emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "fg_richness_mean") 
#  fringing different from all others
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef           0.1872 0.0468 Inf   3.999  0.0004
# Fringing - Forereef 10m       0.2348 0.0519 Inf   4.524  <.0001
# Fringing - Forereef 17m       0.2040 0.0489 Inf   4.173  0.0002
# Backreef - Forereef 10m       0.0477 0.0472 Inf   1.011  0.7432
# Backreef - Forereef 17m       0.0169 0.0439 Inf   0.385  0.9806
# Forereef 10m - Forereef 17m  -0.0308 0.0475 Inf  -0.648  0.9164
#
```

```{r figure 4b plot setup}
rich_stab_emm_fg <- emmip(rich_stab_mod_fg,
                        habitat ~ fg_richness_mean,
                        at = list(fg_richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects_fg <- rich_stab_emm_fg %>%
  left_join(dss_ranges_fg, by = c("habitat")) %>%
  filter(fg_richness_mean >= min_richness_fg & fg_richness_mean <= max_richness_fg)
```

@@

```{r plot figure 4b}
(figure_4b <- 
  ggplot() +
  geom_point(data = fg_stability_quad_level,
             aes(x = fg_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_fg, aes(x = fg_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_stab_effects_fg, aes(x = fg_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "", title = "") +
  trends_theme
         )

@@

filtered_synch_stab_effects <- synch_stab_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
```

@@

filtered_rich_stab_effects_site <- rich_stab_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)
```

@@

#need to get average functional richness through time! 
fg_summary_site_wide_summary <- fg_summary_site_wide %>%
  group_by(habitat, site) %>%
  summarize(avg_richness_fg = mean(fg_richness_site),
            se_richness_fg = sd(fg_richness_site)/sqrt(n()))

#n = 24!

# merge
stability_site_level_fg <- merge(diversity_stability_synchrony_site, fg_summary_site_wide_summary, by = c("site", "habitat"))

# extract relevant ranges for plotting
dss_ranges_site_fg_stab <- stability_site_level_fg %>%
  group_by(habitat) %>%
  summarise(min_richness_fg = min(avg_richness_fg),
            max_richness_fg = max(avg_richness_fg),
            min_stability= min(cover_stability),
            max_stability = max(cover_stability))

rich_stab_mod_site_fg <- glmmTMB(cover_stability~ avg_richness_fg*habitat, family = Gamma(link = "log"), data = stability_site_level_fg)
summary(rich_stab_mod_site_fg)
Anova(rich_stab_mod_site_fg)

@@

hist(residuals(rich_stab_mod_site_fg)) # blocky but fine
plot(residuals(rich_stab_mod_site_fg) ~ fitted(rich_stab_mod_site_fg)) # heteroskedastic
r.squaredGLMM(rich_stab_mod_site_fg) # 0.563
emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "avg_richness_fg") 
# Fringing different from outer 10
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef           0.2552 0.254 Inf   1.005  0.7463
# Fringing - Forereef 10m       1.2611 0.469 Inf   2.686  0.0364
# Fringing - Forereef 17m       0.3543 0.287 Inf   1.235  0.6045
# Backreef - Forereef 10m       1.0059 0.492 Inf   2.044  0.1720
# Backreef - Forereef 17m       0.0991 0.323 Inf   0.307  0.9900
# Forereef 10m - Forereef 17m  -0.9068 0.510 Inf  -1.778  0.2838
```

```{r figure 4e setup}
rich_stab_emm_site_fg <- emmip(rich_stab_mod_site_fg,
                        habitat ~ avg_richness_fg,
                        at = list(avg_richness_fg = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects_site <- rich_stab_emm_site_fg %>%
  left_join(dss_ranges_site_fg_stab, by = c("habitat")) %>%
  filter((avg_richness_fg >= min_richness_fg) & (avg_richness_fg <= max_richness_fg))

@@

```{r figure 4e}
(figure_4e <- 
  ggplot() +
  geom_point(data = stability_site_level_fg,
             aes(x = avg_richness_fg, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_site, aes(x = avg_richness_fg, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = avg_richness_fg, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "Functional Richness", title = "") +
  trends_theme
         )

@@

filtered_synch_stab_effects_site <- synch_stab_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
```

@@

## diversity-synchrony plot level

@@

dss_spatial_ranges <- dss_spatial_2 %>%
  group_by(habitat) %>%
  summarise(min_spatial_synchrony = min(spatial_synchrony, na.rm = TRUE),
            max_spatial_synchrony = max(spatial_synchrony, na.rm = TRUE),
            min_mean_synchrony= min(synchrony_mean, na.rm = TRUE),
            max_mean_synchrony = max(synchrony_mean, na.rm  = TRUE),
            min_mean_stability = min(stability_mean, na.rm = TRUE),
            max_mean_stability = max(stability_mean, na.rm  = TRUE))

## models
```{r model spatial synchrony}
spatial_synchrony_mod <- glmmTMB(cover_stability ~ spatial_synchrony*habitat, family = Gamma("log"), data = dss_spatial_2)
hist(residuals(spatial_synchrony_mod)) # a little skew
plot(residuals(spatial_synchrony_mod) ~ predict(spatial_synchrony_mod))
summary(spatial_synchrony_mod)
Anova(spatial_synchrony_mod)
#                             Chisq Df Pr(>Chisq)    
# spatial_synchrony         16.2914  1  5.431e-05 ***
# habitat                   19.5303  3  0.0002124 ***
# spatial_synchrony:habitat  6.7239  3  0.0812368 .  
emtrends(spatial_synchrony_mod, pairwise ~ habitat, var = "spatial_synchrony") # no sig differences but close
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            0.404 1.765 Inf   0.229  0.9958
# Fringing - Forereef 10m        1.365 1.754 Inf   0.778  0.8643
# Fringing - Forereef 17m        3.112 1.882 Inf   1.654  0.3484
# Backreef - Forereef 10m        0.961 0.857 Inf   1.121  0.6766
# Backreef - Forereef 17m        2.708 1.095 Inf   2.472  0.0644
# Forereef 10m - Forereef 17m    1.747 1.078 Inf   1.621  0.3668
```

filtered_spat_stab_effects <- spat_stab_emm %>%
  left_join(dss_spatial_ranges, by = c("habitat")) %>%
  filter(spatial_synchrony >= min_spatial_synchrony & 
           spatial_synchrony <= max_spatial_synchrony)

(figure_5a <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = spatial_synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Spatial synchrony", title = "") +
  trends_theme
         )

```{r mean synchrony}
mean_synchrony_mod <- glmmTMB(cover_stability ~ synchrony_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
hist(residuals(mean_synchrony_mod)) # chunky
plot(residuals(mean_synchrony_mod) ~ predict(mean_synchrony_mod)) # fine
summary(mean_synchrony_mod)
Anova(mean_synchrony_mod)
#                         Chisq Df Pr(>Chisq)    
# synchrony_mean         32.914  1  9.634e-09 ***
# habitat                18.523  3  0.0003430 ***
# synchrony_mean:habitat 19.062  3  0.0002654 ***
emtrends(mean_synchrony_mod, pairwise ~ habitat, var = "synchrony_mean") # no sig differences but close
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            -3.26 1.503 Inf  -2.166  0.1327
# Fringing - Forereef 10m         2.22 0.995 Inf   2.231  0.1149
# Fringing - Forereef 17m         3.37 1.220 Inf   2.764  0.0291
# Backreef - Forereef 10m         5.48 1.622 Inf   3.377  0.0041
# Backreef - Forereef 17m         6.63 1.769 Inf   3.748  0.0010
# Forereef 10m - Forereef 17m     1.15 1.363 Inf   0.846  0.8322
```

filtered_mean_synch_effects <- mean_synch_emm %>%
  left_join(dss_spatial_ranges, by = c("habitat")) %>%
  filter(synchrony_mean >= min_mean_synchrony & 
           synchrony_mean <= max_mean_synchrony)

(figure_5b <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = synchrony_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_mean_synch_effects, aes(x = synchrony_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_mean_synch_effects, aes(x = synchrony_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Mean synchrony", title = "") +
  trends_theme
         )
```

```{r alpha gamma stability mod plot}
alpha_gamma_stab_mod <- glmmTMB(cover_stability ~ stability_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
summary(alpha_gamma_stab_mod)
Anova(alpha_gamma_stab_mod)
#                          Chisq Df Pr(>Chisq)    
# stability_mean         61.2725  1   4.97e-15 ***
# habitat                12.9323  3   0.004785 ** 
# stability_mean:habitat  9.4832  3   0.023511 *

filtered_ag_effects <- alpha_gamma_emm %>%
  left_join(dss_spatial_ranges, by = c("habitat")) %>%
  filter(stability_mean >= min_mean_stability & 
           stability_mean <= max_mean_stability)

## panel graph
```{r figure 5 panel, fig.height = 6, fig.width=22}
(figure_5 <- ggarrange(figure_5b, figure_5a, alpha_gamma_stability_plot, 
                      ncol = 3, nrow = 1, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      labels = c("a.", "b.", "c."),
                      font.label = list(size = 26, color = "black", face = "plain")))

>>>>

# prep functional group data

@@

diversity_stability_site <- alpha_diversity_site_macro %>%
  dplyr::group_by(site_habitat, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            functional_richness_mean = mean(functional_richness), 
            cover_mean = mean(prop_cover),
            cover_cv = CV(prop_cover),
            cover_stability = 1/CV(prop_cover))

@@

dss_ranges_site <- extract_ranges(diversity_stability_synchrony_site, "habitat",
                                  c("richness_mean", "functional_richness_mean", "cover_stability", "synchrony"))

@@

filtered_rich_stab_effects <- filter_ranges(rich_stab_emm, dss_ranges, "habitat", "richness_mean")
```

@@

## Figure 4b - functional group richness stability

### model
```{r fig4b fg richness stability mod}

rich_stab_mod_fg <- glmmTMB(cover_stability ~ functional_richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)

@@

hist(residuals(rich_stab_mod_fg)) # looks fine
plot(residuals(rich_stab_mod_fg) ~ fitted(rich_stab_mod_fg)) # same as usual 
r.squaredGLMM(rich_stab_mod_fg) # 0.819, 0.828
emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "functional_richness_mean") 
#  fringing different from all others
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef           0.1872 0.0468 Inf   3.999  0.0004
# Fringing - Forereef 10m       0.2348 0.0519 Inf   4.524  <.0001
# Fringing - Forereef 17m       0.2040 0.0489 Inf   4.173  0.0002
# Backreef - Forereef 10m       0.0477 0.0472 Inf   1.011  0.7432
# Backreef - Forereef 17m       0.0169 0.0439 Inf   0.385  0.9806
# Forereef 10m - Forereef 17m  -0.0308 0.0475 Inf  -0.648  0.9164
#
```

```{r figure 4b plot setup}
rich_stab_emm_fg <- emmip(rich_stab_mod_fg,
                        habitat ~ functional_richness_mean,
                        at = list(functional_richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects_fg <- filter_ranges(rich_stab_emm_fg, dss_ranges, "habitat", "functional_richness_mean")
```

@@

```{r plot figure 4b}
(figure_4b <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = functional_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_fg, aes(x = functional_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_stab_effects_fg,
              aes(x = functional_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "", title = "") +
  trends_theme
         )

@@

filtered_synch_stab_effects <- filter_ranges(synch_stab_emm, dss_ranges, "habitat", "synchrony")
```

@@

filtered_rich_stab_effects_site <- filter_ranges(rich_stab_emm_site, dss_ranges_site, "habitat", "richness_mean")
```

@@

rich_stab_mod_site_fg <- glmmTMB(cover_stability~ functional_richness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_site)
summary(rich_stab_mod_site_fg)
Anova(rich_stab_mod_site_fg)

@@

hist(residuals(rich_stab_mod_site_fg)) # blocky but fine
plot(residuals(rich_stab_mod_site_fg) ~ fitted(rich_stab_mod_site_fg)) # heteroskedastic
r.squaredGLMM(rich_stab_mod_site_fg) # 0.563
emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "functional_richness_mean") 
# Fringing different from outer 10
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef           0.2552 0.254 Inf   1.005  0.7463
# Fringing - Forereef 10m       1.2611 0.469 Inf   2.686  0.0364
# Fringing - Forereef 17m       0.3543 0.287 Inf   1.235  0.6045
# Backreef - Forereef 10m       1.0059 0.492 Inf   2.044  0.1720
# Backreef - Forereef 17m       0.0991 0.323 Inf   0.307  0.9900
# Forereef 10m - Forereef 17m  -0.9068 0.510 Inf  -1.778  0.2838
```

```{r figure 4e setup}
rich_stab_emm_site_fg <- emmip(rich_stab_mod_site_fg,
                        habitat ~ functional_richness_mean,
                        at = list(functional_richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects_site <- filter_ranges(rich_stab_emm_site_fg, dss_ranges_site, "habitat", "functional_richness_mean")

@@

```{r figure 4e}
(figure_4e <- 
  ggplot() +
  geom_point(data = diversity_stability_site,
             aes(x = functional_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_site, aes(x = functional_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_stab_effects_site, 
              aes(x = functional_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "Functional Richness", title = "") +
  trends_theme
         )

@@

filtered_synch_stab_effects_site <- filter_ranges(synch_stab_emm_site, dss_ranges_site, "habitat", "synchrony")
```

@@

dss_spatial_ranges <- extract_ranges(dss_spatial_2, "habitat", c("spatial_synchrony", "synchrony_mean", "stability_mean"))

### spatial synchrony
```{r model spatial synchrony}
spatial_synchrony_mod <- glmmTMB(cover_stability ~ spatial_synchrony*habitat, family = Gamma("log"), data = dss_spatial_2)
hist(residuals(spatial_synchrony_mod)) # a little skew
plot(residuals(spatial_synchrony_mod) ~ predict(spatial_synchrony_mod))
summary(spatial_synchrony_mod)
Anova(spatial_synchrony_mod)
#                             Chisq Df Pr(>Chisq)    
# spatial_synchrony         16.2914  1  5.431e-05 ***
# habitat                   19.5303  3  0.0002124 ***
# spatial_synchrony:habitat  6.7239  3  0.0812368 .  
emtrends(spatial_synchrony_mod, pairwise ~ habitat, var = "spatial_synchrony") # no sig differences but close
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            0.404 1.765 Inf   0.229  0.9958
# Fringing - Forereef 10m        1.365 1.754 Inf   0.778  0.8643
# Fringing - Forereef 17m        3.112 1.882 Inf   1.654  0.3484
# Backreef - Forereef 10m        0.961 0.857 Inf   1.121  0.6766
# Backreef - Forereef 17m        2.708 1.095 Inf   2.472  0.0644
# Forereef 10m - Forereef 17m    1.747 1.078 Inf   1.621  0.3668
```

filtered_spat_stab_effects <- filter_ranges(spat_stab_emm, dss_spatial_ranges, "habitat", "spatial_synchrony")

(spatial_synchrony_plot <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = spatial_synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Spatial synchrony", title = "") +
  trends_theme
         )

### mean synchrony
```{r mean synchrony}
mean_synchrony_mod <- glmmTMB(cover_stability ~ synchrony_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
hist(residuals(mean_synchrony_mod)) # chunky
plot(residuals(mean_synchrony_mod) ~ predict(mean_synchrony_mod)) # fine
summary(mean_synchrony_mod)
Anova(mean_synchrony_mod)
#                         Chisq Df Pr(>Chisq)    
# synchrony_mean         32.914  1  9.634e-09 ***
# habitat                18.523  3  0.0003430 ***
# synchrony_mean:habitat 19.062  3  0.0002654 ***
emtrends(mean_synchrony_mod, pairwise ~ habitat, var = "synchrony_mean") # no sig differences but close
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            -3.26 1.503 Inf  -2.166  0.1327
# Fringing - Forereef 10m         2.22 0.995 Inf   2.231  0.1149
# Fringing - Forereef 17m         3.37 1.220 Inf   2.764  0.0291
# Backreef - Forereef 10m         5.48 1.622 Inf   3.377  0.0041
# Backreef - Forereef 17m         6.63 1.769 Inf   3.748  0.0010
# Forereef 10m - Forereef 17m     1.15 1.363 Inf   0.846  0.8322
```

filtered_mean_synch_effects <- filter_ranges(mean_synch_emm, dss_spatial_ranges, "habitat", "synchrony_mean")

(mean_synchrony_plot <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = synchrony_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_mean_synch_effects, aes(x = synchrony_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_mean_synch_effects, aes(x = synchrony_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Mean synchrony", title = "") +
  trends_theme
         )
```

### mean stability plot
```{r alpha gamma stability mod plot}
alpha_gamma_stab_mod <- glmmTMB(cover_stability ~ stability_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
summary(alpha_gamma_stab_mod)
Anova(alpha_gamma_stab_mod)
#                          Chisq Df Pr(>Chisq)    
# stability_mean         61.2725  1   4.97e-15 ***
# habitat                12.9323  3   0.004785 ** 
# stability_mean:habitat  9.4832  3   0.023511 *

filtered_ag_effects <- filter_ranges(alpha_gamma_emm, dss_spatial_ranges, "habitat", "stability_mean")

## panel graph
```{r figure 5 panel, fig.height = 6, fig.width=22}
(figure_5 <- ggarrange(mean_synchrony_plot, alpha_gamma_stability_plot, spatial_synchrony_plot,
                      ncol = 3, nrow = 1, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      labels = c("a.", "b.", "c."),
                      font.label = list(size = 26, color = "black", face = "plain")))

## Trends in richness over time

### plot
```{r figure 2a}
(figure_s1a <- ggplot() +
  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
             position = position_dodge(0.4)) +
  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
                                             colour = habitat), width = 0, size = 1.3, position = position_dodge(0.4)) +
  scale_colour_manual(values = habitat_colours) +
  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
   labs(y = "Macroalgal Richness",
        x = "Year",
        title = "") +
   trends_theme)

# ggsave(filename = "output/figure_s1a.png", figure_s1a, height = 10, width = 14)
```

### plot
```{r plot figure 2b}
(figure_s1b <- ggplot() +
  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
             position = position_dodge(0.4)) +
#  geom_jitter(data = alpha_diversity_site_macro, 
#  aes(x = year, y = richness, colour = habitat, alpha = 0.05), size = 2, width = 0.4, height = 0) +
  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0,  position = position_dodge(0.4)) +
  scale_colour_manual(values = habitat_colours) +
  annotate("segment", x = 2006, xend = 2010, y = 12, yend = 12, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_richness_df$year)) +
   labs(y = "",
        x = "Year",
        title = "") +
  trends_theme)

# ggsave(filename = "output/figure_s1b.png", figure_s1b, height = 10, width = 14)
```

## Figure 2 Full Plot
```{r ggpubr fig2, fig.height = 10, fig.width=22}
fig2_plot_list <- list(figure_s1a, figure_s1b)
(figure_s1 <- ggarrange(figure_s1a, figure_s1b,
                      ncol = 2, nrow = 1, 
                      labels = c("a.", "b."),
                      common.legend = TRUE,
                      legend = "bottom", 
                      font.label = list(size = 24, color = "black", face = "plain")))

# ggsave(filename = "output/figure_s1v4.png", figure_s1, height = 10, width = 22)
```

## diversity-synchrony relationships: plot level

@@
